Using Distributed Information Stores for Patient-Clinician Health Data Sharing

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Storing/representing update

Currently these products are designed to work in isolation, that is, to upload sensor measurements to dedicated web sites set up by each device's manufacturer.  Such sites are poorly suited for the long-term storage for the highly personal activity data for several reasons; first, none of the sites provide any long-term retention or access guarantees of user data (nor HIPAA security guarantees).  Second, it is difficult for a user to gain a unified understanding of their activity across their devices through the disparate interfaces provided by these services, due of the widely different visual and interface representations provided. Finally, many of these visualisations represent user activity in arbitrary, 'user-friendly' units, such as 'Sleep score', 'Fuel', ``Zen level'', which, being not rigorously defined are difficult or impossible to compare across services.

Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.  We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.  In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.

The data collected by patients' devices has the potential to create a large volume of highly structured data. This is in contrast to
EHRs which tend to contain individual records for each clinical session.

To this challenge we bring our contribution of a common ontology for recording and encompassing patient activities, which we have begun to
connect to SNOMED-CT, in order to enable semantic interoperability with existing systems.

>>>>

Currently these products are designed to work in isolation, that is, to upload sensor measurements to dedicated web sites set up by each device's manufacturer.  Such sites are poorly suited for the long-term storage for the highly personal activity data for several reasons; first, none of the sites provide any long-term retention or access guarantees of user data (nor HIPAA security guarantees).  Second, it is difficult for a user to gain a unified understanding of their activity across their devices through the disparate interfaces provided by these services, due of the widely different visual and interface representations provided. Finally, many of these visualisations represent user activity in arbitrary, ``user-friendly'' units, such as ``Sleep score'', ``Fuel'', ``Zen level'', which, being not rigorously defined are difficult or impossible to compare across services.

Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.  We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single representation kept with their EHR.  In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.

To this challenge we bring our contribution of a common ontology for recording and encompassing patient activities, which we have begun to
connect to SNOMED-CT, in order to enable semantic interoperability with existing systems, and provide a common vocabulary across devices.

The data collected by patients' devices has the potential to create a large volume of highly structured data. Some sensors
are collecting minute-by-minute data values form multiple sensors 24 hours a day, whereas 
EHRs tend to contain individual records for each clinical session. Therefore, existing methods of storing EHRs may
prove to be unsuitable for the large volumes of data from activity diaries. However, integrating with existing EHRs
would enable clinicians and patients to view the data in context.
